Generating Random Logic Programs Using Constraint Programming

### 2020-09-12 paper/paper.tex
Bring back the appendix

\renewcommand\fbox{\fcolorbox{red}{white}}
\newcommand{\hilight}[1]{\setlength{\fboxsep}{1pt}\colorbox{lightgray}{#1}}

@@

\title{Generating Random Logic Programs Using Constraint Programming}
\author{Paulius Dilkas\inst{1} \and Vaishak Belle\inst{1,2}$^{(\textrm{\Letter})}$}
\authorrunning{P. Dilkas and V. Belle}
\institute{University of Edinburgh, Edinburgh, UK
  \email{\{p.dilkas@sms.,vaishak@\}ed.ac.uk} \and Alan Turing Institute, London,
  UK}

@@

\begin{figure}[t]
  \centering
  \begin{minipage}{.49\textwidth}
    \centering
    \begin{tikzpicture}
      \node[draw] (parent) at (0, 0.5) {$\mathsf{parent}$};
      \node[draw] (mother) at (0, -0.5) {$\mathsf{mother}$};
      \node[draw] (sibling) at (2, 0.5) {$\mathsf{sibling}$};
      \node[draw] (father) at (2, -0.5) {$\mathsf{father}$};
      \draw[-{Stealth[scale=1.5]}] (parent) edge node[above] {$+$} (sibling);
      \draw[-{Stealth[scale=1.5]}] (parent) edge node[above] {$+$} (father);
      \draw[-{Stealth[scale=1.5]}] (mother) edge node[below] {$-$} (father);
    \end{tikzpicture}%
    \captionof{figure}{The predicate dependency graph of the program from
      \protect{\cref{ex:program}}. Positive edges are labelled with `$+$', and
      negative edges with `$-$'.}
    \label{fig:predicate_dependencies}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \centering
    \captionof{table}{Types of (potential) dependencies of a predicate
      $\mathsf{P}$ based on the number of undetermined edges on the path from
      the dependency to $\mathsf{P}$}
    \label{tbl:dependencies}
    \begin{tabular}{clc}
      \toprule
      Edges & Name & Notation \\
      \midrule
      $0$ & determined & $\Determined(p)$ \\
      $1$ & almost determined & $\AlmostDetermined(p, s, t)$ \\
      $>1$ & undetermined & $\Undetermined(p)$ \\
      \bottomrule
    \end{tabular}
  \end{minipage}
\end{figure}

@@

>>>>

\renewcommand\fbox{\fcolorbox{red}{white}}
\newcommand{\hilight}[1]{\setlength{\fboxsep}{1pt}\colorbox{lightgray}{#1}}
\newcommand{\hlitem}{\stepcounter{enumi}\item[\hilight{\theenumi}]}

@@

\title{Generating Random Logic Programs Using Constraint Programming}
\author{Paulius Dilkas\inst{1} \and Vaishak Belle\inst{1,2}$^{(\textrm{\Letter})}$}
\authorrunning{P. Dilkas and V. Belle}
\institute{University of Edinburgh, Edinburgh, UK \\
  \email{p.dilkas@sms.ed.ac.uk, vaishak@ed.ac.uk} \and Alan Turing Institute,
  London, UK}

@@

\begin{figure}[t]
  \centering
  \begin{minipage}{.49\textwidth}
    \centering
    \begin{tikzpicture}
      \node[draw] (parent) at (0, 0.5) {$\mathsf{parent}$};
      \node[draw] (mother) at (0, -0.5) {$\mathsf{mother}$};
      \node[draw] (sibling) at (2, 0.5) {$\mathsf{sibling}$};
      \node[draw] (father) at (2, -0.5) {$\mathsf{father}$};
      \draw[-{Stealth[scale=1.5]}] (parent) edge node[above] {$+$} (sibling);
      \draw[-{Stealth[scale=1.5]}] (parent) edge node[above] {$+$} (father);
      \draw[-{Stealth[scale=1.5]}] (mother) edge node[below] {$-$} (father);
    \end{tikzpicture}%
    \captionof{figure}{The predicate dependency graph of the program from
      \protect{\cref{ex:program}}. Positive edges are labelled with `$+$', and
      negative edges with `$-$'.}
    \label{fig:predicate_dependencies}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \centering
    \captionof{table}{Types of (potential) dependencies of a predicate
      $\mathsf{P}$ based on the number of undetermined edges on the path from
      the dependency to $\mathsf{P}$}
    \label{tbl:dependencies}
    \begin{tabular}{clc}
      \toprule
      Edges & Name & Notation \\
      \midrule
      $0$ & Determined & $\Determined(p)$ \\
      $1$ & Almost determined & $\AlmostDetermined(p, s, t)$ \\
      $>1$ & Undetermined & $\Undetermined(p)$ \\
      \bottomrule
    \end{tabular}
  \end{minipage}
\end{figure}

@@

\newpage
\appendix
\section{Example Programs}

In this appendix, we provide examples of probabilistic logic programs generated
by various combinations of parameters. In all cases, we use
\[
  \{ 0.1, 0.2, \dots, 0.9, 1, 1, 1, 1, 1\}
\]
as the multiset of probabilities. Each clause is written on a separate line and
ends with a full stop. The head and the body of each clause are separated with
\texttt{:-} (instead of $\gets$). The probability of each clause is prepended to
the clause, using \texttt{::} as a separator. Probabilities equal to one and
empty bodies of clauses can be omitted. Conjunction, disjunction, and negation
are denoted by commas, semicolons, and `\texttt{\textbackslash+}', respectively.
Parentheses are used to demonstrate precedence, although many of them are
redundant.

By setting $\predicates{} = [\texttt{p}]$, $\arities{} = [1]$, $\variables{} =
\{ \texttt{X} \}$, $\constants{} = \emptyset$, $\maxNumNodes{} = 4$, and
$\maxNumClauses{} = 1$, we get fifteen one-line programs, six of which are
without negative cycles (as highlighted below). Only the last program has no
cycles at all.

\begin{enumerate}[label=\arabic*.]
\item
\begin{verbatim}
0.5 :: p(X) :- (\+(p(X))), (p(X)).
\end{verbatim}
\item
\begin{verbatim}
0.8 :: p(X) :- (\+(p(X))); (p(X)).
\end{verbatim}
\hlitem
\begin{verbatim}
0.8 :: p(X) :- (p(X)); (p(X)).
\end{verbatim}
\hlitem
\begin{verbatim}
0.7 :: p(X) :- (p(X)), (p(X)).
\end{verbatim}
\item
\begin{verbatim}
0.6 :: p(X) :- (p(X)), (\+(p(X))).
\end{verbatim}
\item
\begin{verbatim}
p(X) :- (p(X)); (\+(p(X))).
\end{verbatim}
\hlitem
\begin{verbatim}
0.1 :: p(X) :- (p(X)); (p(X)); (p(X)).
\end{verbatim}
\hlitem
\begin{verbatim}
0.8 :: p(X) :- (p(X)), (p(X)), (p(X)).
\end{verbatim}
\item
\begin{verbatim}
p(X) :- \+(p(X)).
\end{verbatim}
\item
\begin{verbatim}
0.1 :: p(X) :- \+(\+(p(X))).
\end{verbatim}
\item
\begin{verbatim}
p(X) :- \+((p(X)); (p(X))).
\end{verbatim}
\item
\begin{verbatim}
0.4 :: p(X) :- \+((p(X)), (p(X))).
\end{verbatim}
\item
\begin{verbatim}
0.4 :: p(X) :- \+(\+(\+(p(X)))).
\end{verbatim}
\hlitem
\begin{verbatim}
0.7 :: p(X) :- p(X).
\end{verbatim}
\hlitem
\begin{verbatim}
p(X).
\end{verbatim}
\end{enumerate}

Note that:
\begin{itemize}
\item A program such as Program~14, because of its cyclic definition, defines a
  predicate that has probability zero across all constants. This can more easily
  be seen as solving equation $0.7x = x$.
\item Programs 10 and 14 are not equivalent (i.e., double negation does not
  cancel out) because Program 10 has a negative cycle and is thus considered to
  be ill-defined.
\end{itemize}

To demonstrate variable symmetry reduction in action, we set $\predicates{} =
[\texttt{p}]$, $\arities{} = [3]$, $\variables{} = \{\texttt{X}, \texttt{Y},
\texttt{Z} \}$, $\constants{} = \emptyset$, $\maxNumNodes{} = 1$,
$\maxNumClauses{} = 1$, and forbid all cycles. This gives us the following five
programs:

\begin{itemize}
\item
\begin{verbatim}
0.8 :: p(Z, Z, Z).
\end{verbatim}
\item
\begin{verbatim}
p(Y, Y, Z).
\end{verbatim}
\item
\begin{verbatim}
p(Y, Z, Z).
\end{verbatim}
\item
\begin{verbatim}
p(Y, Z, Y).
\end{verbatim}
\item
\begin{verbatim}
0.1 :: p(X, Y, Z).
\end{verbatim}
\end{itemize}

This is one of many possible programs with $\predicates{} = [\texttt{p},
\texttt{q}, \texttt{r}]$, $\arities{} = [1, 2, 3]$, $\variables{} =
\{\texttt{X}, \texttt{Y}, \texttt{Z} \}$, $\constants{} = \{ \texttt{a},
\texttt{b}, \texttt{c} \}$, $\maxNumNodes{} = 5$, $\maxNumClauses{} = 5$, and
without negative cycles:

\begin{verbatim}
p(b) :- \+((q(a, b)), (q(X, Y)), (q(Z, X))).
0.4 :: q(X, X) :- \+(r(Y, Z, a)).
q(X, a) :- r(Y, Y, Z).
q(X, a) :- r(Y, b, Z).
r(Y, b, Z).
\end{verbatim}

Finally, we set
$\predicates{} = [\texttt{p}, \texttt{q}, \texttt{r}]$, $\arities{} = [1, 1,
1]$, $\variables{} = \emptyset$, $\constants{} = \{ \texttt{a} \}$,
$\maxNumNodes{} = 3$, $\maxNumClauses{} = 3$, forbid negative cycles, and
constrain predicates \texttt{p} and \texttt{q} to be independent. The resulting
search space contains thousands of programs such as:

\begin{itemize}
\item
\begin{verbatim}
0.5 :: p(a) :- (p(a)); (p(a)).
0.2 :: q(a) :- (q(a)), (q(a)).
0.4 :: r(a) :- \+(q(a)).
\end{verbatim}
\item
\begin{verbatim}
p(a) :- p(a).
0.5 :: q(a) :- (r(a)); (q(a)).
r(a) :- (r(a)); (r(a)).
\end{verbatim}
\item
\begin{verbatim}
p(a) :- (p(a)); (p(a)).
0.6 :: q(a) :- q(a).
0.7 :: r(a) :- \+(q(a)).
\end{verbatim}
\end{itemize}
